---
title: "PA1_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
 
```{r loaddata echo=TRUE}
mydata<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
 
```{r 1 echo=TRUE}
total_steps_day<-tapply(mydata$steps,mydata$date,sum,na.rm=TRUE)
mean(total_steps_day)
median(total_steps_day)
par(mfrow=c(1,1))
hist(total_steps_day,xlab = "toal number of steps each day", main = "Histogram of the total number of \n steps each day")
```
The mean value is 9354.23, and the median is 10395.


## What is the average daily activity pattern?

```{r 2 echo=TRUE}
time_serious_steps<-aggregate(mydata$steps,list(mydata$interval),mean,na.rm=TRUE)
par(mfrow=c(1,1)) 
plot(time_serious_steps[,1],time_serious_steps[,2], type="l", xlab = "interval",ylab="mean steps")
time_serious_steps[which.max(time_serious_steps[,2]),1]
```
The 835 interval contains the maximum number of steps.
## Imputing missing values
```{r 3 echo=TRUE}
sum(is.na(mydata$steps))

# use the mean of that interval to repalce NA values
mydata_raplaceNA<-mydata
mean_interval<-aggregate(mydata$steps,list(mydata$interval),mean,na.rm=TRUE)
mydata_raplaceNA$steps[is.na(mydata$steps)]<-mean_interval[which(mydata$interval[is.na(mydata$steps)]==mean_interval[,1]),2]

total_steps_day_noNA<-tapply(mydata_raplaceNA$steps,mydata_raplaceNA$date,sum,na.rm=TRUE)
par(mfrow=c(1,1))
hist(total_steps_day_noNA, xlab = "toal number of steps each day", main = "Histogram of the total number of \n steps each day (imputed missing values)")
mean(total_steps_day_noNA)
median(total_steps_day_noNA)
```
The total number of missing values in the dataset is 2304. The new data without NA is called mydata_replaceNA.The mean and median values of the new data are different from those two values of the old data.

## Are there differences in activity patterns between weekdays and weekends?
```{r 4 echo=TRUE}
library(dplyr)
mydata_raplaceNA$status<- ifelse(weekdays(as.POSIXct(mydata_raplaceNA$date),abbreviate = T)=="Sat"|weekdays(as.POSIXct(mydata_raplaceNA$date),abbreviate = T)=="Sun","weekend","weekday")

time_serious_steps_noNA<-aggregate(mydata_raplaceNA$steps,list(mydata_raplaceNA$interval,mydata_raplaceNA$status),mean, na.rm=TRUE)

 library(lattice)
par(mfrow=c(1,1))
xyplot(x~Group.1|Group.2, data=time_serious_steps_noNA, type="l", layout=c(1,2),ylab = "Number of steps",xlab = "Interval")

 
```


 